---
title: "Geographic Bias in Paper Titles"
author: "C. Sean Burns, Md. Anwar Islam"
output: html_document
params:
  dataset: "titles"
---

```{r include = FALSE}
library(tidyverse)
library(dplyr)
library(conflicted)
library(ggplot2)
library(rstatix)
library(effsize)
library(knitr)

# Import data
titles <- read.csv("~/workspace/geo-titles/data/titlesMain.csv")
```

# Introduction 

Writing a title for a paper is complicated and hard. Authors must write a title
that succinctly, yet informatively, captures the main topic of their paper.
Since titles function within the knowledge creation process, titles reflect
disciplinary identity and expectations [@HGX4FHTA#milojević17]. In some
disciplines, like math, titles focus on succinctness. Yet in other areas, like
the social sciences, titles stress informativeness. This leads to various
conventions, such as the use of colons in titles. Titles with colons (compound
titles) are less succinct and potentially more informative. They move from
making general statements (left of colon) to specific statements (right of
colon). The specificity ranges from descriptive to declarative statements.
Succinct titles may focus on providing topical information only. Compound
titles add information about the method, research design, results, or
conclusion of a study [@8VIPNBL7#li19].

Title informativeness can be helpful when searching literature for a topic. The
informativeness of a title can be a function of the words it contains and its
length or word count. Theoretically, the more unique words in a title, the more
likely a searcher might locate the work in an online search. In the economics
literature, longer titles receive more citations ("the informative effect")
than shorter, succinct titles ("the succinct effect"), but this is only true
after the year 2000. @NV6JFP8H#guo18 attributes this to the rise of online
searching, where retrieval technology is based on relevance algorithms that
index keywords in various bibliographic fields. @8VIPNBL7#li19 also found that
title length started to increase during this time frame, but defined title
informativeness not solely based on word count but on the proportion of content
words (e.g., nouns, verbs, adverbs) to function words (e.g., pronouns,
prepositions, conjunctions). Titles with a higher ratio of content words are
more informative. This may be complicated though if those function words are
overly specific, obscure, or non-normative in some way [@AKWVYLLV#fox15,
@NLQ23JTG#thelwall17].

If a title poorly captures the information about the content of a paper, then
papers may be overlooked by searchers even if the papers are relevant to
searchers. Alternatively, if title information is framed in such a way as to
seem non-applicable, even if the paper is relevant to a searcher, then such
papers may be overlooked. For example, papers with titles that ask questions
have been shown to receive more downloads than papers with other title types
but fewer citations [@JUCB9NIU#jamali11], or simply fewer citations
[@U67Y7URB#paiva12]. This suggests that the information captured by a
question-type title is not as informative, or perhaps more trendy, than other
title types and therefore results in a relevance mismatch (i.e., it seemed
relevant but was not) in online searches or browsing. However, a disciplinary
effect exists. Papers with titles that ask questions are cited more in the
computer science literature [@IQEEVG95#fiala21] and no citation effects were
found for question-type titles in an ecology journal [@AKWVYLLV#fox15].

As mentioned, the use of compound titles (titles with colons, hyphens, dashes)
has grown in recent decades, especially in some research areas or disciplines.
@8VIPNBL7#li19 outlines three types of compound titles that capture specific
semantic content. These are titles that describe the topic and the method or
design, titles that describe the topic and the results, and titles that
describe the topic and the conclusion of a study. In addition to these
variations, some authors add geographical names to compound titles, which may
not add key semantic information about a paper's topic [@9966FJNU#kou18]. A
number of studies have shown that papers with titles that contain certain types
of highly specific or obscure content, like genus or species information
[@AKWVYLLV#fox15], or specific geographical place names, receive fewer
citations [ @LQN8UTGB#abramo16, @C65KJSGU#jacques10, @U67Y7URB#paiva12,
@L4QDBN7X#costello19, @RRLX7MYL#moradi16, @NLQ23JTG#thelwall17]. The common
explanation is that this kind of taxonomic or geographic specificity is
extraneous to the study, and that it reduces the generality and thus the appeal
of the results reported in these papers to searchers or potential audiences
[@AKWVYLLV#fox15, @NLQ23JTG#thelwall17]. However, pointing to more disciplinary
differences, specificity might be appreciated in some disciplines, like
entomology, where titles with specific genus and species names or geographic
names have been shown to have greater impact [@GL3HSUBY#murphy19].

@LQN8UTGB#abramo16 and others [@AKWVYLLV#fox15, @NLQ23JTG#thelwall17] have
reasoned that overly specific terms or words in titles, especially those naming
geographic entities, tend to receive fewer citations because searchers
reviewing these titles do not find them relevant even if the topic is relevant.
@LQN8UTGB#abramo16 suggested, for papers with geographical names in titles,
that "studies conducted at the country level would typically be less appealing
that [sic] those dealing with the same subjects at the broader level. The
researcher [or potential reader] could suspect that certain results would be
influenced by country-specific traits, and therefore be difficult to
generalize" (p. 13). However, papers without specific geographical names in the
titles might still be limited to specific geographical areas, and this raises
questions not only about why some authors include geographical information in
titles, but also why some authors do not. 

Only one study, that we could find, has shown different patterns in how
geographic names are used in titles among researchers around the world. Based
on a five year period of papers published in the CHI Conference Proceedings,
@9966FJNU#kou18 found differences in whether titles contain geographic names
depending on the population studied. They wrote that "studies conducted with
non-Western populations are significantly more likely to highlight study
contexts in titles and throughout the text," and that "studies of Western
countries are significantly more likely to lack mention of the studied
countries not only in titles, but also throughout the text of the papers"
[@9966FJNU#kou18, p. 2]. They reported that these findings held even after
controlling for the authors' countries. They also found that when country names
were added to titles, they were often added at the end of a title, which rarely
suggested "deep semantic connections between the preceding ideas in the titles
and the countries" (p. 8). Overall, their study suggests a geographical bias
in, at least the CHI literature, to normalize findings based on Western
populations and to exoticize findings based on non-Western populations. 

Like @9966FJNU#kou18, we are also interested how authors use geographic names
in titles. However, @9966FJNU#kou18 apply a simple binary classification of
countries into Western and non-Western countries based on work by
@46QMHZS8#huntington11. However, like @B5Z3GY8X#burns17, we use the Human
Development Index (HDI) [@UBTFXCKV#unitednationsdevelopmentprogramme] to
identify patterns in how countries are named in paper titles since we believe
this should offer are more nuanced view of the data. Based on this, we propose
the following research questions:

1. Which countries are more likely to be named in paper titles? 
2. Are countries named in titles more likely to have lower HDI values?
3. Do geographical names in titles provide important semantic content about
   some aspect of the study?

# Materials and Methods

Our data is focused on library and information science journal titles,
as broadly categorized and listed by Scimago’s Journal and Country Rank (2021).
We limited our set to the 61 journal titles ranked in the first quartile.
We downloaded bibliographic records for articles published from 2018 to 2020
in these titles from Scopus.
For example, our Scopus query for the journal
*International Journal of Information Management* was:

```
SRCTITLE ( "International Journal of Information Management" )  AND  (
LIMIT-TO ( PUBYEAR ,  2020 )  OR  LIMIT-TO ( PUBYEAR ,  2019 )  OR  LIMIT-TO
( PUBYEAR ,  2018 ) )  AND  ( LIMIT-TO ( DOCTYPE ,  "ar" ) )
```

To prepare the data, we imported it into R using the bibliometrix package.
Our title data set includes 13,145 records across the 61 journals.
To extract place names from the article titles,
we used the *spacy* natural language processing Python library.
This identified 1,239 (9.4%) article titles with place names.

We used the 2021-2022 Human Development Index (HDI) from the
United Nations Development Programme that matched
the place name or named region in the title.
The HDI is a composite index that attempts to capture three dimensions:
life expectancy index, education index, and GINI index.
Most place names used in titles matched place names in the HDI;
however, some modifications were needed.
Some but very few article titles had place names,
like Taiwan or North Korea,
that reference nations that do not have an HDI.
These were marked as NA.
If a place did not have an HDI but was part of a broader collective,
like Scotland or Puerto Rico,
we chose the HDI for the broader collective
(i.e., the United Kingdom or the United States).
The Human Development Index also includes region level values.
When article titles referenced Africa,
we used the HDI for Sub-Saharan Africa, and
when article titles referenced Europe,
we used the HDI for "Europe and Central Asia"
When multiple place names appeared in the title (*n* = 86),
we averaged the HDI scores and created a
separate variable to tag these as averaged.
The *spacy* library was able to identify nation names
for places like U.S. states.
Thus, if an article title mentioned a place like "Alaska",
then United States was used for the nation and HDI variables.
Alternatively, if a title included a term like *american*,
the *spacy* library was able to reference this as the *United States*.
Therefore, titles may include a number of place name variations, but
going forward, we refer to the canonical name for the nation that is
referenced in the title.

Finally, we added SCImago Journal Rank (SJR) scores
for each of the 61 journal titles in the data set,
We use the SJR to control for citation effects across articles.
SJR scores ranged from 4.585 to 0.528.

# Results

```{r include = FALSE}
# data prep
# subset of data that does not include titles with place names
without_nations <- dplyr::filter(titles, is.na(nation))
without_nations <- dplyr::select(without_nations, Citations, SJR)
# subset of data that does includes titles with place names
with_nations <- dplyr::filter(titles, nation != "NA" & HDI != "NA")
with_nations <- dplyr::select(with_nations, Citations, SJR, nation, HDI)
```

There were 202 unique place names,
or place name combinations,
identified in the 1236 articles with place names in titles and with
HDI scores.
China appeared in titles most frequently,
followed by Spain, the U.S.A., the United Kingdom, and India.
Table 1 lists the 20 most frequently referenced place names in titles.

## Countries and Titles

Which countries are more likely to be named in paper titles? 

```{r echo = FALSE}
my_tab <- titles %>%
  dplyr::filter(nation != "NA" & HDI != "NA") %>%
  dplyr::select(nation) %>% table
my_tab <- my_tab %>% as.data.frame %>%
  dplyr::arrange(desc(Freq)) %>% head(20)
```

```{r echo = FALSE}
my_tab %>% slice_head(n = 20) %>%
  kable(caption = "Table 1. Top 20 most frequent nations referred to in article titles")
```

```{r echo = FALSE}
rm(my_tab)
```


## Countries and Citations

Are titles with country names more likely to have lower citations?

A Wilcoxon rank-sum test was conducted
to compare the median citation counts of
papers with places names in titles (Mdn = 6)
to papers without place names in titles (Mdn = 7).
The test revealed a significant difference
between the two groups
(W = 6863554, p = 0.0001, *n* with place names = 1236,
*n* without place names = 11898),
indicating that papers with place names
in titles had received significantly
lower citations than those without.
The null hypothesis that there is
no difference between the two groups
was rejected in favor of the alternative
hypothesis that papers without place names in titles
receive more citations
than papers with place names in titles.
These results suggest that papers without place names
in titles may receive more citations, in general,
than papers with place names in titles.
However, Cliff's delta was calculated to assess the
magnitude of the difference in scores between 
papers with place names in titles
to papers without place names in titles.
The analysis revealed a negligible effect size
in favor of papers without place names in titles
(delta = -0.0666, 95% CI [-0.100, -0.033]).

```{r include = FALSE}
# citation differences between nation and non-nation data subsets
summary(with_nations$Citations)
sd(with_nations$Citations)
summary(without_nations$Citations)
sd(without_nations$Citations)
wilcox.test(with_nations$Citations, without_nations$Citations)
cliff.delta(with_nations$Citations, without_nations$Citations)
```

```{r echo = FALSE}
citations_df <- data.frame(
  group = c(rep("With Nations", length(with_nations$Citations)),
            rep("Without Nations", length(without_nations$Citations))),
  score = c(log(with_nations$Citations + 1),
            log(without_nations$Citations + 1)
))
ggplot(citations_df, aes(x = group, y = score)) +
  geom_boxplot() +
  labs(x = "Groups", y = "Citations (log + 1 transformed)",
       title = "Comparison of Citations between Papers
       With Nations Named in Titles and Without") +
  theme_minimal()
```

## Countries and Impact

Are titles with country names more likely to be published in lower impact journals?

Papers with place names in titles publish in lower impact journals,
(*mdn* = 0.93)
based on SJR,
than papers without place names in titles
(*mdn* = 1.07).
A Wilcoxon rank-sum test was conducted
to compare the median scores.
The test revealed a significant difference
between the two groups
(W = 6358560, p < 0.0001, *n* with place names = 1236,
*n* without place names = 11898),
indicating that papers with place names
in titles published in journal titles
with significantly lower impact scores than
papers without place names in titles.
The null hypothesis that there is
no difference between the two groups
was rejected in favor of the alternative
hypothesis that papers without place names in titles
publish in lower impact journals.
However, Cliff's delta was calculated to assess the
magnitude of the difference between 
journal impact scores.
The analysis revealed a negligible effect size
showing that papers without place names in titles
tend to publish in journals with lower impact scores
than papers with place names in titles
(delta = -0.1352, 95% CI [-0.168, 0.102]).

```{r include = FALSE}
# SJR differences between nation and non-nation data subsets
summary(with_nations$SJR)
sd(with_nations$SJR)
summary(without_nations$SJR)
sd(without_nations$SJR)
wilcox.test(with_nations$SJR, without_nations$SJR)
cliff.delta(with_nations$SJR, without_nations$SJR)
```

```{r echo = FALSE}
sjr_df<- data.frame(
  group = c(rep("With Nations", length(with_nations$SJR)),
            rep("Without Nations", length(without_nations$SJR))),
  score = c(with_nations$SJR, without_nations$SJR)
)
ggplot(sjr_df, aes(x = group, y = score)) +
  geom_boxplot() +
  labs(x = "Groups", y = "Journal Impact (SJR)",
       title = "Comparison of Journal Impact Scores
       With Nations Named in Titles and Without") +
  theme_minimal()
```

## Journals and Titles with Countries

What journals have the most titles with place names?  
Are these journals more likely to have lower SJRs?

We suspected that journals with lower impact scores might publish more papers
with place names in titles. Therefore we partitioned the list of articles
with places names in titles into three subsets: set1 contains articles in
journals less than the median SJR for the set,
set2 contains titles with greater than the median SJR,
and set3 contains papers in journals equal to SJR median.

We found that papers with countries named in
titles were significantly less likely to be published in journals
with above median SJR scores (*n* = 491; *Mdn* = 1.503)
compared to journals with below median
SJR scores (*n* = 592; *Mdn* = 0.756).
A portion of papers with countries named in titles were published
in journals with SJRs equal to the median (*n* = 153; *Mdn* = 0.929)
A Wilcoxon ran-sum test was conducted to compare the median SJRs
between these the two groups below and above the median.
The test revealed a significant difference
(W = 290672; *p* < 0.0001).

```{r include = FALSE}
SJR_med             <- with_nations %>% mutate(med = median(SJR))
less_than_median    <- SJR_med %>% dplyr::filter(SJR < med)
greater_than_median <- SJR_med %>% dplyr::filter(SJR > med)
equal_to_median     <- SJR_med %>% dplyr::filter(SJR == med)
wilcox.test(greater_than_median$SJR, less_than_median$SJR)
```

```{r echo = FALSE}
sjr_df_mdn <- data.frame(
  group = c(rep("Less than Mdn",    length(less_than_median$SJR)),
            rep("Greater than Mdn", length(greater_than_median$SJR))),
  score = c(less_than_median$SJR, greater_than_median$SJR)
)
ggplot(sjr_df_mdn, aes(x = group, y = score)) +
  geom_boxplot() +
  labs(x = "Groups", y = "Journal Impact (SJR)",
       title = "Articles with titles more often appear
       in journals with lower impact scores") +
  theme_minimal()
```

## Countries and HDIs

```{r echo = FALSE}
ggplot(with_nations, aes(x = HDI, y = log(Citations + 1))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "HDI", y = "Citations (log + 1)", title = "HDI to Citations") +
  theme_light()

ggplot(with_nations, aes(x = SJR, y = log(Citations + 1))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "SJR", y = "Citations (log + 1)", title = "SJR to Citations") +
  theme_light()
```

```{r echo = FALSE}
my_tab <- with_nations %>% arrange(desc(Citations)) %>%
  select(nation, Citations, SJR, HDI) %>%
  slice_head(n = 20)
```

```{r echo = FALSE}
my_tab %>% slice_head(n = 20) %>%
  kable(caption = "Table 2. Papers with nations named or referred to with highest citations")
rm(my_tab)
```


# Discussion

Papers naming or referring to countries in titles
receive fewer citations and are published in lower
ranking journals than papers that do not name or refer
to countries in titles.
Although Cliff's delta reported negligible effect sizes
for citation results,
we think that a cumulative effect might be in play.
That is, if any particular author or journal tends
to publish papers with place names in titles,
then the citation scores for those authors or journals
might receive fewer citations depending on the
frequency of papers with place names in titles.

Limitations: it could be that the effects are more pronounced
when sampling from lower impact journals.

Results could be different with a longer citation window.

# Conclusion

# References
