---
title: "Place Names in Article Titles Correspond to Fewer Citations"
author: "C. Sean Burns, Md. Anwar Islam"
output:
  html_document:
    theme: cerulean
    keep_md: yes
  word_document: default
  pdf_document: default
params:
  dataset: titles
editor_options:
  markdown:
    wrap: 72
  chunk_output_type: console
zotero: Titles
bibliography: Titles.bib
csl: scientometrics.csl
---

```{r load-libraries, include = FALSE}
library(tidyverse)
library(conflicted)
library(rstatix)
library(effsize)
library(knitr)
library(ggridges)
library(rbbt)

# Import data
titles <- read.csv("~/workspace/geo-titles/data/titlesMain.csv")
titles$HDI <- round(titles$HDI, 3)
```

# Introduction

A good article title informs and can serve to attract readers.
Authors must write titles that succinctly
capture the main topic of their paper.
Since titles function within the knowledge creation process,
titles reflect disciplinary identity and expectations [@milojevic_length_2017].
In some disciplines, like math, titles focus on succinctness.
Yet in other areas, like the social sciences,
titles stress informativeness.
In the medical sciences,
the use of questions in titles has increased substantially
since the 1960s,
representing perhaps institutional pressures
to publish faster [@ballScholarlyCommunicationTransition2009]
@ballScholarlyCommunicationTransition2009 also found geographic trends
in the use of question marks in titles among some disciplines.

The need to market papers, express disciplinary identity, provide information,
and other factors leads to various naming conventions,
such as the use of colons in titles.
Titles with colons (compound titles) are less succinct and
potentially more informative.
They move from making general statements (left of colon)
to specific statements (right of colon).
The specificity ranges from descriptive to declarative statements.
Succinct titles may focus on providing topical information.
Compound titles may add information about the method,
research design, results, or conclusion of a study [@li_evolution_2019].

Title informativeness can be helpful when searching literature for a topic.
The informativeness of a title can be a function of the
words it contains and its length or word count.
In the economics literature,
longer titles receive more citations ("the informative effect")
than shorter, succinct titles ("the succinct effect"),
but this is only true after the year 2000.
@guo_succinct_2018 attributes this to the rise of online searching,
where retrieval technology is based on relevance algorithms that index
keywords in various bibliographic fields.
@li_evolution_2019 found that title length started
to increase during this time frame, but
defined title informativeness not solely based on word count but on the
proportion of content words (e.g., nouns, verbs, adverbs) to function
words (e.g., pronouns, prepositions, conjunctions).
Titles with a higher ratio of content words are more informative.
This may be complicated though if those function words are overly specific,
obscure, or non-normative in some way [@fox_relationship_2015;
@thelwall_avoiding_2017].

If a title poorly captures the information about the content of a paper,
then papers may be overlooked by searchers even if the papers are
relevant to them. Alternatively, if title information is framed in
such a way as to seem non-applicable, even if the paper is relevant to a
searcher, then such papers may be overlooked. For example, papers with
titles that ask questions have been shown to receive more downloads but
fewer citations than
papers with other title types [@jamali_article_2011; @paiva_articles_2012].
This suggests that the information captured by a question-type
title is informative but trendy, not as informative, or
that the authors are less certain of the findings.
However, a disciplinary effect exists. Papers with
titles that ask questions are cited more in the computer science
literature [@fiala_are_2021] and no citation effects were found for
question-type titles in an ecology journal [@fox_relationship_2015]

As mentioned, the use of compound titles (titles with colons, hyphens,
dashes) has grown in recent decades, especially in some research areas
or disciplines. @li_evolution_2019 outlines three types of compound titles
that capture specific semantic content. These are titles that describe
the topic and the method or design, titles that describe the topic and
the results, and titles that describe the topic and the conclusion of a
study. In addition to these variations, some authors add geographical
names to compound titles, which may not add key semantic information
about a paper's topic [@kou_politics_2018]. A number of studies have shown
that papers with titles that contain certain types of highly specific or
obscure content, like genus or species information [@fox_relationship_2015], or
specific geographical place names, receive fewer citations
[@abramo_effect_2016; @costello_are_2019; @jacques_impact_2010; @moradi_analysis_2016; @paiva_articles_2012; @thelwall_avoiding_2017].
The common explanation is that this kind of taxonomic or geographic
specificity is extraneous to the study, and that it reduces the
generality and thus the appeal of the results reported in these papers
to searchers or potential audiences [@fox_relationship_2015;
@thelwall_avoiding_2017]. However, pointing to more disciplinary
differences, specificity might be appreciated in some disciplines, like
entomology, where titles with specific genus and species names or
geographic names have been shown to have greater impact [@murphy_does_2019].

@abramo_effect_2016 and others [@fox_relationship_2015; @thelwall_avoiding_2017]
have reasoned that overly specific terms or words in titles, especially
those naming geographic entities, tend to receive fewer citations
because searchers reviewing these titles do not find them relevant even
if the topic is relevant. @abramo_effect_2016 suggested, for papers with
geographical names in titles, that "studies conducted at the country
level would typically be less appealing that [sic] those dealing with
the same subjects at the broader level. The researcher [or potential
reader] could suspect that certain results would be influenced by
country-specific traits, and therefore be difficult to generalize" (p.
13). However, papers without specific geographical names in the titles
might still be limited to specific geographical areas, and this raises
questions not only about why some authors include geographical
information in titles but also why some authors do not.

Only one study, that we could find, has shown different patterns in how
geographic names are used in titles among researchers around the world.
Based on a five year period of papers published in the CHI Conference
Proceedings, @kou_politics_2018 found differences in whether titles contain
geographic names depending on the population studied. They wrote that
"studies conducted with non-Western populations are significantly more
likely to highlight study contexts in titles and throughout the text,"
and that "studies of Western countries are significantly more likely to
lack mention of the studied countries not only in titles, but also
throughout the text of the papers" [@kou_politics_2018, p. 2]. They
reported that these findings held even after controlling for the
authors' countries. They also found that when country names were added
to titles, they were often added at the end of a title, which rarely
suggested "deep semantic connections between the preceding ideas in the
titles and the countries" (p. 8). Overall, their study suggests a
geographical bias in, at least the CHI literature, to normalize findings
based on Western populations and to exoticize findings based on
non-Western populations.

Like @kou_politics_2018, we are also interested how authors use geographic
names in titles. However, @kou_politics_2018 apply a simple binary
classification of countries into Western and non-Western countries based
on work by @huntington_clash_2011. However, like @burns_language_2017, we
use the Human Development Index (HDI)
[@united_nations_development_programme_human_2020], an index that measures
a nation's level of health, education, and standard of living, to identify patterns in
how countries are named in paper titles. We believe this 
offer a more nuanced view of the data. The UN classifies HDI scores into
four categories: very high (\>= 0.8), high (\>= 0.7), medium (\>= 0.55),
and low (\<= 0.549). Based on this, we ask the following questions:

1. Do articles with titles that name countries receive fewer citations than articles with titles without countries?
2. Are there social and economic reasons that explain citation differences?

# Materials and Methods

```{r count_data, include = FALSE, warning = FALSE}
n_total <- titles %>% dplyr::select(Citations) %>% summarize(count = n())
n_nations <- titles %>% dplyr::select(nation) %>%
  dplyr::filter(!is.na(nation)) %>% summarize(count = n())
n_nations_per <- round((n_nations / n_total) * 100, 2)
```

Our data is focused on articles published in
library and information science (LIS) journals,
as broadly categorized and listed by Scimago's Scientific Journal Ranking (SJR)
and the Scimago Journal and Country Rank (2021) data.
We limited our set of articles to the 61 journals ranked
in the first quartile of the LIS category.
We conducted Scopus queries for each journal
in this list and downloaded bibliographic data for articles
published in these journals from 2018 to 2020.
We retrieved `r n_total` article records across the 61 journals.
For example, our Scopus query for the journal
*International Journal of Information Management* was:

```         
SRCTITLE ( "International Journal of Information Management" )  AND  (
LIMIT-TO ( PUBYEAR ,  2020 )  OR  LIMIT-TO ( PUBYEAR ,  2019 )  OR  LIMIT-TO
( PUBYEAR ,  2018 ) )  AND  ( LIMIT-TO ( DOCTYPE ,  "ar" ) )
```

We combined and imported the bibliographic records into
R using the bibliometrix package [@aria_bibliometrix_2017].
We used the *spaCy* [@honnibal_spacy_2022]
natural language processing Python library to identify 
country names from article titles.
This identified `r n_nations` (`r n_nations_per`%)
article titles with place names.
We used the 2021-2022 Human Development Index (HDI) from the
United Nations (UN) Development Programme
that matched the country in the titles.
The HDI summarizes three indices:
a life expectancy index, an education index, and
GINI index for measuring income inequality.

```{r no_HDIs, include = FALSE, warning = FALSE}
no_HDIs <- titles %>% dplyr::select(nation, HDI) %>%
  dplyr::filter(!is.na(nation)) %>%
  dplyr::filter(is.na(HDI)) %>%
  summarize(count = n())
nations_no_HDIs <- n_nations - no_HDIs 
averaged_HDIs <- titles %>% dplyr::select(Averaged) %>%
  dplyr::filter(grepl('Y', Averaged)) %>% summarize(count = n())
```

Most places named in titles matched names in the HDI.
However, some articles had titles with country names that for political reasons
do not have an HDI (e.g., Taiwan and North Korea).
These were marked as NA in the HDI variable.
This reduced titles with country names and HDI scores to `r nations_no_HDIs` records.
If a place did not have an HDI but was part of a broader collective
(e.g., Scotland or Puerto Rico),
we chose the HDI for the broader political authority
(i.e., the United Kingdom or the United States, respectively).
The HDI includes region level values.
When article titles referenced Africa,
we used the HDI for "Sub-Saharan Africa", and
when article titles referenced Europe,
we used the HDI for "Europe and Central Asia".
When multiple country names appeared in the title (*n* = `r averaged_HDIs`),
we averaged the HDI scores for the named places.

The Python *spaCy* library was able to identify country names for constituent locales,
such as specific states in the U.S.
Thus, if an article title mentioned a place like 'Alaska',
then 'United States' was used for the country and HDI variables.
Alternatively, if a title included a term like *american*,
the *spacy* library was able to reference this as the *United States*.
Therefore, titles may include a number of place name variations.
If so, we refer to the canonical name for the country that
is referenced or inferred in the title.

We were interested in the location of the country name in each article title.
The *spaCy* natural processing library detected the location of
866 country names in article titles.
Country positions were scored on a scale of 0 to 1.0,
with countries appearing as the last word in a title
receiving a score of 1.0.
In cases where multiple countries were named in the title,
the score was based on the first instance.

[Add spacy accuracy statement here.]

Finally, we added SCImago Journal Rank (SJR) scores for each of the 61
journal titles in the data set.
We use the SJR to control for citation effects across articles.
SJR scores ranged from 0.528 to 4.585.
A SJR score below 1.0 indicates below average citations compared to all
journals in *Scopus*.

# Results

[Write intro to Results]

```{r divide-data, include = FALSE, warning = FALSE}
# data prep
# subset of data that does not include titles with place names
without_nations <- dplyr::filter(titles, is.na(nation))
without_nations <- dplyr::select(without_nations, Citations, SJR)
# subset of data that does includes titles with place names
with_nations <- dplyr::filter(titles, nation != "NA" & HDI != "NA")
with_nations <- dplyr::select(with_nations, Citations, SJR, nation, HDI)
```

## Countries and Journal Impact

```{r tabularize-nations, include = FALSE}
median(with_nations$HDI)
titles %>% dplyr::select(nation) %>% table %>% as.data.frame()
with_nations %>% dplyr::select(nation) %>% table %>% as.data.frame()
```

```{r uniq-nations, include = FALSE}
uniq_count <- titles %>% dplyr::select(nation) %>%
  dplyr::filter(nation != "NA") %>%
  unique() %>% count()
```

```{r top-20-nations, echo = FALSE}
my_tab <- with_nations %>%
  dplyr::filter(nation != "NA" & HDI != "NA") %>%
  dplyr::select(nation, HDI) %>% table
wn_df <- my_tab %>% as.data.frame %>%
  dplyr::arrange(desc(Freq)) %>% dplyr::filter(Freq > 0)
my_tab_20 <- my_tab %>% as.data.frame %>%
  dplyr::arrange(desc(Freq)) %>% slice_head(n = 20)
```

There were `r uniq_count`
country names or country name combinations
identified in the `r length(with_nations$nation)` articles with
country names in titles and with HDI scores.
Out of these, `r dim(my_tab)[1]`
had HDI scores assigned to them.
China appeared in titles most frequently, followed by Spain, the U.S.A.,
the United Kingdom, and India.
The median HDI for the observed countries is
`r median(with_nations$HDI)`.
Therefore, most identified countries are classified as having
very high human development index scores
(*min* = `r min(with_nations$HDI)`,
*m* = `r mean(with_nations$HDI)`
*max* = `r max(with_nations$HDI)`).
Table 1 lists the `r dim(my_tab_20)[1]` most frequently referenced
country names in titles along with their respective HDI scores.

```{r top-20-nations-table, echo = FALSE}
my_tab_20 %>% 
  kable(caption = "Table 1. Top 20 countries in article titles")
rm(my_tab)
```

### Journals with Articles with Country Names in Titles Receiver Fewer Citations

Journals that receive SJR scores above 1.0 indicate journals that
receive above average citations compared to all journals in *Scopus*.
The average SJR for all the journals in our data was slightly
above the *Scopus* average (*mdn* = `r median(titles$SJR)`).
We sought to test whether country names in article titles had
an impact on journal citations.
We divided the data into two groups:
articles with country names referenced in titles and
articles without country names referenced in titles.
We compared the SJR of the journals between these two groups.
Even though there are journal titles that belong to both groups,
our goal was to observe the effect that country names
have on overall journal citations.

We found that the inclusion of country names in article titles had a
negative effect on citations to journals.
The average SJR for journals that published
articles that did not include references to country names was 
higher (*mdn* = `r median(without_nations$SJR)`) than the SJR average.
However, the average SJR for journals that published
articles that did include references to countries was lower than the
*Scopus* average and lower than journals with articles without
references to countries (*mdn* = `r median(with_nations$SJR)`).
Overall, this suggests that journals that tend to
publish articles with titles containing references
to countries receive fewer citations than journals that tend to
publish articles to do not reference countries.
Table 2 reports the journal titles that most frequently
publish articles that reference country names in article titles.

```{r get-summaries, include = FALSE}
summary(titles$SJR)
summary(with_nations$SJR)
summary(without_nations$SJR)
```

```{r get-SJRs, include = FALSE}
titles %>% dplyr::select(Publication) %>% table %>% as.data.frame() %>%
  dplyr::arrange(desc(Freq)) %>% slice_head(n = 20)
with_nations %>% dplyr::select(SJR) %>% table %>% as.data.frame() %>%
  dplyr::arrange(desc(Freq)) %>% slice_head(n = 20)
```

```{r get-table-2, echo = FALSE}
my_tab <- titles %>%
  dplyr::filter(nation != "NA" & HDI != "NA") %>%
  dplyr::select(Publication, SJR) %>% table
my_tab <- my_tab %>% as.data.frame %>%
  dplyr::arrange(desc(Freq)) %>% slice_head(n = 20)
my_tab %>%
  kable(caption = "Table 2. Top 20 most frequent journal titles with countries mentioned in titles")
```

```{r get-wt-test-1, include = FALSE}
# SJR differences between country and non-country data subsets
summary(with_nations$SJR)
sd(with_nations$SJR)
summary(without_nations$SJR)
sd(without_nations$SJR)
nations_wt <- wilcox.test(with_nations$SJR, without_nations$SJR)
nations_cd <- cliff.delta(with_nations$SJR, without_nations$SJR)
```

We compared the average SJR of journals with articles with country names
in titles to the SJR of journals without country names in articles
using a Wilcoxon rank-sum test.
The test revealed a significant difference 
(W = `r nations_wt$statistic`,
p \< `r round(nations_wt$p.value, 3)`,
*n* with place names = `r dim(with_nations)[1]`
*n* without place names = `r dim(without_nations)[1]`).
However, even though articles without countries in titles
published in lower impact journals,
the overall effect size was small
(Cliff's delta = `r nations_cd$estimate`,
95% CI [`r nations_cd$conf.int[1]`, `r nations_cd$conf.int[2]`]).

```{r group-data, include = FALSE}
sjr_df<- data.frame(
  group = c(rep("With Countries", length(with_nations$SJR)),
            rep("Without Countries", length(without_nations$SJR))),
  score = c(with_nations$SJR, without_nations$SJR)
)
```

```{r fig-1, fig.cap = "Fig. 1: Comparison of Journal Impacts Scores", echo = FALSE}
ggplot(sjr_df, aes(x = group, y = score)) +
  geom_boxplot() +
  labs(x = "Groups", y = "Journal Impact (SJR)") +
  theme_minimal()
```

### Lower impact journals publish more articles with country names in titles

We suspected that journals with lower impact scores publish more
articles with place names in titles.
Therefore we partitioned the list of articles with places names in
titles into three subsets:
set1 contains articles in journals with less than the median SJR,
set2 contains articles in journals with greater than the median SJR, and
set3 contains articles in journals equal to SJR median.

```{r compare-median-SJRs, include = FALSE}
SJR_med             <- with_nations %>% mutate(med = median(SJR))
less_than_median    <- SJR_med %>% dplyr::filter(SJR < med)
greater_than_median <- SJR_med %>% dplyr::filter(SJR > med)
equal_to_median     <- SJR_med %>% dplyr::filter(SJR == med)
grlt_wt <- wilcox.test(greater_than_median$SJR, less_than_median$SJR)
```

Although some articles with countries named in titles were
published in journals with SJRs equal to the median
(*n* = `r length(equal_to_median$SJR)`;
*mdn* = `r median(equal_to_median$SJR)`),
we found that articles with country names in titles were
less likely to appear in journals with above median SJR scores
(*n* = `r length(greater_than_median$SJR)`;
*mdn* = `r median(greater_than_median$SJR)`)
and more likley to appear in journals with below median SJR scores
(*n* = `r length(less_than_median$SJR)`;
*mdn* = `r median(less_than_median$SJR)`).

We applied a Wilcoxon ran-sum test to compare the below and
above median groups. The test revealed a significant difference
(W = `r grlt_wt$statistic`;
*p* \< `r round(grlt_wt$p.value, 3)`),
indicating that journals with higher SJR scores
publish fewer articles with country names in titles.

```{r publication-sorting, echo = FALSE}
SJR_journals_with <- titles %>% arrange(Publication) %>%
  dplyr::filter(nation != "NA" & HDI != "NA") %>%
  dplyr::select(Publication, SJR)
```

```{r fig-2a, echo = FALSE}
testdf1 <- unique(SJR_journals_with)
testdf2 <- table(SJR_journals_with$Publication)
testdf2 <- as.data.frame(testdf2)
testdf1$Freq <- testdf2$Freq
testdf1 <- testdf1 %>% dplyr::arrange(desc(Freq))
```

```{r get-non-title-data, echo = FALSE}
SJR_journals_without <- titles %>% 
  dplyr::filter(is.na(nation)) %>%
  dplyr::select(Publication, SJR) %>%
  arrange(Publication)
```

```{r fig-2b, echo = FALSE}
testdf3 <- unique(SJR_journals_without)
testdf4 <- table(SJR_journals_without$Publication)
testdf4 <- as.data.frame(testdf4)
testdf3$Freq <- testdf4$Freq
testdf3 <- testdf3 %>% dplyr::arrange(desc(Freq))
```

```{r fig-2, echo = FALSE, fig.cap = "Fig. 2: Articles and SJR"}
testdf1$Country <- "with_country"
testdf3$Country <- "without_country"
combined_df <- bind_rows(testdf1, testdf3)
ggplot(combined_df, aes(x = SJR, y = Freq, size = Freq, color = Country)) + geom_point(alpha = 0.5) + theme_light() + scale_color_grey()
```

## Countries and Article Impact

```{r get-wt-test-2, include = FALSE}
# citation differences between country and non-country data subsets
summary(with_nations$Citations)
sd(with_nations$Citations)
summary(without_nations$Citations)
sd(without_nations$Citations)
citations_wt <- wilcox.test(with_nations$Citations, without_nations$Citations)
citations_cd <- cliff.delta(with_nations$Citations, without_nations$Citations)
```

We found that article titles with country names are more likely to have
lower citation counts than article titles without country names.
A Wilcoxon rank-sum test was conducted to compare the median citation counts. 
Articles with countries referenced in titles received fewer citations
(*mdn* = `r median(with_nations$Citations)`)
than articles without references to countries in titles
(*mdn* = `r median(without_nations$Citations)`).
The test revealed a significant difference between the two groups
(W = `r citations_wt$statistic`,
*p* = `r round(citations_wt$p.value, 3)`,
*n* with place names = `r length(with_nations$Citations)`,
*n* without place names = `r length(without_nations$Citations)`).
These results suggest that papers without place names in titles received
more citations, in general, than papers with place names in titles.
However, the effect size was small.
We used Cliff's delta to assess the magnitude of the difference in
scores between papers with place names in titles to papers without place
names in titles.
The analysis revealed a negligible effect size in favor of papers
without place names in titles
(delta = `r citations_cd$estimate`,
95% CI [`r citations_cd$conf.int`]).

```{r group-citation-data, echo = FALSE}
citations_df <- data.frame(
  group = c(rep("With Countries", length(with_nations$Citations)),
            rep("Without Countries", length(without_nations$Citations))),
  score = c(log(with_nations$Citations + 1),
            log(without_nations$Citations + 1)
))
```

```{r fig-4, echo = FALSE, fig.cap = "Fig. 4: Comparison of Citations"}
ggplot(citations_df, aes(x = group, y = score)) +
  geom_boxplot() +
  labs(x = "Groups", y = "Citations (log + 1 transformed)") +
  theme_minimal()
```

## Countries and HDIs

```{r get-regression, include = FALSE}
less_HDI <- with_nations %>%
  dplyr::select(HDI) %>% dplyr::filter(HDI < median(HDI)) %>% tally()
great_HDI <- with_nations %>%
  dplyr::select(HDI) %>% dplyr::filter(HDI > median(HDI)) %>% tally()
fit.1 <- lm(log(with_nations$Citations + 1) ~ with_nations$HDI)
fit.cit2hdi <- summary(fit.1)
fstatistic <- fit.cit2hdi$fstatistic
p_value_fstatistic <- pf(fstatistic[1], fstatistic[2],
                         fstatistic[3], lower.tail = FALSE)
```

Countries with below median HDI scores
(*n* = `r less_HDI`)
were as likely to
appear in article titles as countries with higher HDI scores
(*n* = `r great_HDI`).
A linear regression analysis was conducted to examine the
relationship between HDI and citations for articles that reference
country names in titles.
The regression equation was not significant
(F-statistic: `r round(fit.cit2hdi$fstatistic[1], 3)`,
*p* = `r round(p_value_fstatistic, 3)`,
R-squared \< `r round(fit.cit2hdi$r.squared, 3)`),
and the results indicated that the HDI
was not a significant predictor of citation counts
(t-value = `r round(fit.cit2hdi$coefficients[2,3], 3)`,
*p* = `r round(fit.cit2hdi$coefficients[2,4], 3)`,
*b* = `r round(fit.cit2hdi$coefficients[2], 3)`).
We therefore found no significant effect on whether the
HDI of the named country in an article title had an effect on
the number of citations the article received.

```{r fig-5, echo = FALSE, message = FALSE, fig.cap = "Fig. 5: HDI to Citations"}
ggplot(with_nations, aes(x = HDI, y = log(Citations + 1))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "HDI", y = "Citations (log + 1)") +
  theme_light()
```

```{r prep-table-3, echo = FALSE}
my_tab <- with_nations %>% arrange(desc(Citations)) %>%
  select(nation, Citations, SJR, HDI) %>%
  slice_head(n = 20)
```

```{r table-3, echo = FALSE}
my_tab %>%
  kable(caption = "Table 3. Papers with countries named or referred to with highest citations")
rm(my_tab)
```

## Country Naming and Semantic Value

```{r title-positions, include = FALSE}
# Import data with place name positions
titles_place_pos <- read.csv("~/workspace/geo-titles/data/titles-with-place-positions.csv")
```

We examined where the location of countries appeared in article titles
under the assumption that countries appearing near the end of a title
provided little informational value to the idea expressed in the first
part of the title.

```{r get-position-correlation, include = FALSE}
summary(titles_place_pos$places)
titles_place_pos_citations <- titles_place_pos %>%
  dplyr::select(Citations, places) %>%
  dplyr::filter(places != "NA")
places_cor <- cor(titles_place_pos_citations$Citations,
                  titles_place_pos_citations$places, method = "spearman")
```

On average, countries tended to be referenced near the end of a title
(*m* = `r round(mean(titles_place_pos$places, na.rm = TRUE), 3)`,
*mdn* = `r median(titles_place_pos$places, na.rm = TRUE)`),
suggesting that most articles that contain
names of countries in the data do so without adding much semantic
information to the title.
A Spearman's rank correlation was computed to compare the location of
the country name in a title and the citations received to the articles
with country names in titles. We found no relationship
(*r* = `r round(places_cor, 3)`).

```{r fig-6, echo = FALSE, message = FALSE, fig.cap = "Fig. 6: Title Place Position and Citations"}
ggplot(titles_place_pos_citations, aes(x = places,
                                       y = log(Citations + 1))) +
  geom_point() +
  theme_minimal()
```

```{r fig-7, fig.cap = "Fig. 7: Publications with countries listed in articles", echo = FALSE, message = FALSE}
pubs <- titles %>% dplyr::select(Publication, HDI, SJR, nation) %>%
  dplyr::filter(nation != "NA" & HDI != "NA")
ggplot(pubs, aes(x = HDI, y = Publication, fill = SJR)) +
  geom_density_ridges_gradient()
```

# Discussion

This study brought for us a wide set of findings which addressed the
research questions that we set for the study.
Initially, the present study has examined whether
the presence of a country name in the title reduce the
chances of getting more citations.
These findings are discussed.
Papers naming or referring to countries in titles receive fewer
citations and are published in lower ranking journals than papers that
do not name or refer to countries in titles.
In a broad sense, the scientific publications containing
with a country name receive lower impact values e.g.,
the impact factor (IF) of the publishing journrals.
However, under the different subject categories
the result can be different [@abramoEffectCountryName2016].
Although Cliff's delta
reported negligible effect sizes for citation results, we think that a
cumulative effect might be in play. That is, if any particular author or
journal tends to publish papers with place names in titles, then the
citation scores for those authors or journals might receive fewer
citations depending on the frequency of papers with place names in titles.
This result provides general confirmation of our study.
This might be reason that putting the country name in the title
limits the generality of the research findings, and
country-level title is less appealing to the global researchers.
While it is read less, this leads to get fewer citations.

However, under the countries and HDIs,
this can be different in the context of non-western countries as well.
For example, a researcher from a researcher from developing countries has
the intuition to read the USA-based authors’ papers to know the context and
understand the research in his or her similar research areas.
This helps and increases the chances of the developing country researchers
to read some good quality papers.
So, country name in the title and if it is written by western authors might
get the citations by the developing country researchers.
Country names adding at the end of the title e.g. right part of the title
do not add much semantic information.
For example, 'Information seeking behaviours of the graduate students:
A study on the students from Bangladesh' where the subtitle does not
add much information or value to the article if researchers prepare the title
with the left part of the article.
However, this might be different with some disciplines and articles where the
lengthy titles and occurrences of words in the title matters.
However, these practices are not fixed and have changed over times.
While the title is an important factor to the researchers,
there need to be a balance between informativeness and attractiveness. 

Although papers that reference country names in titles received fewer
citations and journals that frequently include articles that refer to 
named countries had lower SJR scores,
we could find no relationship between the development level of a
named country with the number of citations.
We could also find no relationship between the location of the country
name in the title with the article's citations.
Overall, although articles that reference country names in their titles
receive fewer citations and journals have lower SJR scores,
there does not seem to be a geographical bias against specific countries.
We believe that the citation disadvantage associated with lower citation
counts to articles that reference named countries is primarily a result
of perhaps unnecessary specificity that naming countries adds to the titles and
that makes the articles seem less relevant to others.
While the country name in the article’s title refer to less citation and
placed in the lower ranked journals,
we recommend not to use  country name in the title.
Even having a country name in the title may lead to bias among the researchers.  

Limitations: It could be that the effects are more pronounced when
sampling from lower impact journals in some other disciplines.
Here, we only considered the LIS journals.
A future study with the journals from other disciplines can vetted the present
result that we got from this study.
Even, the results could be different with a longer citation window.
We also did not control for other factors,
such if a title contains a colon, a question, humor.

# Conclusion


# References
