---
title: "Place Names in Article Titles Correspond to Fewer Citations"
author: "C. Sean Burns, Md. Anwar Islam"
output:
  html_document:
    theme: cerulean
    keep_md: yes
  word_document: default
  pdf_document: default
params:
  dataset: titles
editor_options:
  markdown:
    wrap: 72
  chunk_output_type: console
zotero: Titles
bibliography: Titles.bib
csl: scientometrics.csl
---

```{r load-libraries, include = FALSE}
library(tidyverse)
library(conflicted)
library(rstatix)
library(effsize)
library(knitr)
library(ggridges)
library(rbbt)

# Import data
titles <- read.csv("~/workspace/geo-titles/data/titlesMain.csv")
titles$HDI <- round(titles$HDI, 3)
```

# Introduction

A good article title informs and can serve to attract readers.
Authors must write titles that succinctly capture the main topic of their paper.
Since titles function within the knowledge creation process,
titles reflect disciplinary identity and expectations [@milojevic_length_2017].
In some disciplines,
like math,
titles focus on succinctness.
Yet in other areas,
like the social sciences,
titles stress informativeness.
In the medical sciences,
the use of questions in titles has increased substantially since the 1960s,
showing also geographical trends,
representing perhaps institutional pressures to publish faster [@ball2009].

The need to market papers,
to express disciplinary identity,
to provide information,
and other factors leads to various naming conventions,
such as the use of colons in titles.
Titles with colons (compound titles) are less succinct and potentially more informative.
They move from making general statements (left of colon)
to specific statements (right of colon).
The specificity ranges from descriptive to declarative statements.
Succinct titles may focus on providing topical information.
Compound titles may add information about the method,
research design, results, or conclusion of a study [@li_evolution_2019].

Title informativeness can be helpful when searching literature for a topic.
The informativeness of a title can be a function of the words
it contains and its length or word count.
In the economics literature,
longer titles receive more citations ("the informative effect")
than succinct titles ("the succinct effect"),
but this is only true after the year 2000.
@guo_succinct_2018 attributes this to the rise of online searching,
where retrieval technology is based on relevance algorithms
that index keywords in various bibliographic fields.
@li_evolution_2019 found that title length started to increase during this time frame,
but defined title informativeness not solely based on word count
but on the proportion of content words
(e.g., nouns, verbs, adverbs) to function words
(e.g., pronouns, prepositions, conjunctions).
Titles with a higher ratio of content words are more informative.
This may be complicated though if function words are overly specific,
obscure, or non-normative in some way
[@fox_relationship_2015; @thelwall_avoiding_2017].

If a title poorly captures the information about the content of a paper,
then papers may be overlooked by searchers
even if the papers are relevant to them.
Alternatively, if title information is framed in such a way as to seem non-applicable,
even if the paper is relevant to a searcher,
then such papers may be overlooked.
For example, papers with titles that ask questions
have been shown to receive more downloads but fewer citations
than papers with other title types [@jamali_article_2011; @paiva_articles_2012].
This suggests that the information captured by
a question-type title is trendy, not informative, or
that the authors are less certain of the findings.
However, a disciplinary effect exists.
Papers with titles that ask questions are cited more
in the computer science literature [@fiala_are_2021]
and no citation effects were found for question-type
titles in an ecology journal [@fox_relationship_2015].

As mentioned, the use of compound titles
(titles with colons, hyphens, dashes) has grown in recent decades,
especially in some research areas or disciplines.
@li_evolution_2019 outlines three types of compound titles
that capture specific semantic content.
These are titles that describe the topic and the method or design,
titles that describe the topic and the results, and
titles that describe the topic and the conclusion of a study.
Additionally, some authors add geographical names to compound titles,
which may not add key semantic information about a paper's topic [@kou_politics_2018].
Studies have shown that papers with titles
that contain certain types of highly specific or obscure content,
like genus or species information [@fox_relationship_2015],
or specific geographical place names,
receive fewer citations
[@abramo_effect_2016; @costello_are_2019; @jacques_impact_2010;
@moradi_analysis_2016; @paiva_articles_2012; @thelwall_avoiding_2017].
The common explanation is that this kind of taxonomic or geographic specificity
is extraneous to the study, and
that it reduces the generality and
thus the appeal of the results reported in these papers to searchers or
potential audiences [@fox_relationship_2015; @thelwall_avoiding_2017].
However, pointing to more disciplinary differences,
specificity might be appreciated in some disciplines,
like entomology,
where titles with specific genus and species names or geographic names
have been shown to have greater impact [@murphy_does_2019].

@abramo_effect_2016 and others
[@fox_relationship_2015; @thelwall_avoiding_2017]
have reasoned that overly specific terms or words in titles,
especially those naming geographic entities,
tend to receive fewer citations because searchers reviewing these titles do not
find them relevant even if the topic is relevant.
@abramo_effect_2016 suggested,
for papers with geographical names in titles, that
"studies conducted at the country level would typically be less appealing
that [sic] those dealing with the same subjects at the broader level.
The researcher [or potential reader] could suspect that certain results
would be influenced by country-specific traits,
and therefore be difficult to generalize" (p. 13).
However, papers without specific geographical names in the titles
might still be limited to specific geographical areas,
and this raises questions not only about why some authors
include geographical information in titles
but also why some authors do not.

Only one study,
that we could find,
has shown different patterns in how geographic names are used
in titles among researchers around the world.
Based on a five year period of papers published in the CHI Conference Proceedings,
@kou_politics_2018 found that
"studies conducted with non-Western populations are significantly
more likely to highlight study contexts in titles and throughout the text,"
and that
"studies of Western countries are significantly more likely
to lack mention of the studied countries not only in titles,
but also throughout the text of the papers" [@kou_politics_2018, p. 2].
They also found that when country names were added to titles,
they were often added at the end of a title,
which suggested weak
"semantic connections between the preceding ideas in the titles and the countries" (p. 8).
Overall, their study suggests a geographical bias in the CHI literature
to normalize findings based on Western populations and
to exoticize findings based on non-Western populations.

Like @kou_politics_2018,
we are interested how authors use geographic names in titles.
However, @kou_politics_2018 apply a simple binary classification
of countries into Western and non-Western countries based on work by @huntington_clash_2011.
In accordance with @burns_language_2017,
we use the Human Development Index (HDI) [@united_nations_human_2023],
an index that measures a nation's level of
health, education, and standard of living,
to identify patterns in how countries are named in paper titles.
We believe this offer a more nuanced view of the data.
Based on this, we ask the following questions:

RQ1: Does the inclusion of geographical names in the titles
of journal articles impact their citation counts,
when controlling for journal impact factor and Human Development Index?

Our research question implies two sub questions:

1. Is there a correlation between the position of the geographical name in the
   article title and the number of citations?
2. Does the Human Development Index of the nation mentioned in the title
   correlate with the number of citations?

In order to answer these questions, we propose the following hypotheses:

* Main Hypothesis: Journal articles with geographical names in their titles are
  cited less frequently than those without geographical names.
    * Sub-Hypothesis 1: The closer the geographical name is to the beginning of the title, the fewer the citations.
    * Sub-Hypothesis 2: Articles mentioning nations with a higher Human Development Index in their titles are cited more frequently.
    * Sub-Hypothesis 3: The impact of geographical names in titles on citation counts is moderated by the impact factor of the journal.
    * Sub-Hypothesis 4: Journals with a high frequency of articles with geographical names in titles have lower impact factors.
    * Sub-Hypothesis 5: Are there any trends with where certain nations appear in the titles?
    * Sub-Hypothesis 6: Journals may be more or less likely to publish articles that mention nations with higher or lower HDIs.

# Materials and Methods

```{r count_data, include = FALSE, warning = FALSE}
n_total <- titles %>% dplyr::select(Citations) %>% summarize(count = n())
n_nations <- titles %>% dplyr::select(nation) %>%
  dplyr::filter(!is.na(nation)) %>% summarize(count = n())
n_nations_per <- round((n_nations / n_total) * 100, 2)
```

Our data is focused on articles published in
library and information science (LIS) journals,
as broadly categorized and listed by Scimago's Scientific Journal Ranking (SJR) and
the Scimago Journal and Country Rank (2021) data.
We limited our set of articles to the 61 journals
ranked in the first quartile of the LIS category.
We conducted Scopus queries for each journal in this list
and downloaded bibliographic data for articles published
in these journals from 2018 to 2020.
We retrieved `r n_total` article records across the 61 journals.
For example, our Scopus query for the journal
*International Journal of Information Management* was:

```         
SRCTITLE ( "International Journal of Information Management" )  AND  (
LIMIT-TO ( PUBYEAR ,  2020 )  OR  LIMIT-TO ( PUBYEAR ,  2019 )  OR  LIMIT-TO
( PUBYEAR ,  2018 ) )  AND  ( LIMIT-TO ( DOCTYPE ,  "ar" ) )
```

We combined and imported the bibliographic records into R
using the bibliometrix package [@aria_bibliometrix_2017].
We used the Python *spaCy* [@honnibal_spacy_2022] natural language processing library
to identify country names from article titles.
We manually reviewed and revised the data after locating
false positives or negatives.
This resulted in  `r n_nations` (`r n_nations_per`%)
article titles with place names.
We added the 2021-2022 Human Development Index (HDI) from the
United Nations (UN) Development Programme that matched the country in the titles.
The HDI summarizes three indices:
a life expectancy index, an education index, and GINI index for measuring income inequality.

```{r no_HDIs, include = FALSE, warning = FALSE}
no_HDIs <- titles %>% dplyr::select(nation, HDI) %>%
  dplyr::filter(!is.na(nation)) %>%
  dplyr::filter(is.na(HDI)) %>%
  summarize(count = n())
nations_no_HDIs <- n_nations - no_HDIs 
averaged_HDIs <- titles %>% dplyr::select(Averaged) %>%
  dplyr::filter(grepl('Y', Averaged)) %>% summarize(count = n())
mean_HDI <- titles %>% dplyr::select(HDI) %>%
  summarize(mean = mean(HDI, na.rm=TRUE)) %>% round(3)
median_HDI <- titles %>% dplyr::select(HDI) %>%
  summarize(mean = median(HDI, na.rm=TRUE)) %>% round(3)
regions <- titles %>% dplyr::select(CountryRegionMulti) %>%
  dplyr::filter(grepl('R', CountryRegionMulti)) %>%
  summarize(count = n())
countries <- titles %>% dplyr::select(CountryRegionMulti) %>%
  dplyr::filter(grepl('C', CountryRegionMulti)) %>%
  summarize(count = n())
multi_nations <- titles %>% dplyr::select(CountryRegionMulti) %>%
  dplyr::filter(grepl('M', CountryRegionMulti)) %>%
  summarize(count = n())
multi_HDI_mean <- titles %>% dplyr::select(HDI, CountryRegionMulti) %>%
  dplyr::filter(grepl('M', CountryRegionMulti)) %>%
  summarize(mean(HDI, na.rm = TRUE)) %>% round(3)
multi_HDI_median <- titles %>% dplyr::select(HDI, CountryRegionMulti) %>%
  dplyr::filter(grepl('M', CountryRegionMulti)) %>%
  summarize(median(HDI, na.rm = TRUE)) %>% round(3)
sjr_unique_mean <- titles %>% dplyr::select(SJR, Publication) %>%
  unique() %>% dplyr::summarize(mean(SJR, na.rm = TRUE)) %>% round(3)
srj_unique_median <- titles %>% dplyr::select(SJR, Publication) %>%
  unique() %>% dplyr::summarize(median(SJR, na.rm = TRUE)) %>% round(3)
```

The Python *spaCy* library was able to identify country names for constituent locales,
such as specific states in the U.S.
Thus, if an article title mentioned a place like "Alaska",
then we used "United States"  for the nation and HDI variables.
Alternatively, if a title included a term like "american",
the *spacy* library was able to reference this as the *United States*.
Therefore, when titles included place name variations,
we referred to the canonical name for the nation referenced or inferred in the title.

We were interested in the location of the place name in each article title.
We wrote a Python script,
calling the *spaCy* natural processing library,
to detect the position of the nation names in article titles.
We manually reviewed all titles to address false positives and negatives.
The position of places in article titles was scored on a scale of 0 to 1.0,
with nations appearing as the last word in a title
receiving a score approaching 1.0.
Titles that did not contain a reference to a geographical region scored a zero.
In cases where multiple countries were named in the title,
the score was based on the first instance.

Most places named in titles matched names in the HDI.
However, some articles had titles with nation names
that for political reasons do not have an HDI (e.g., Taiwan and North Korea).
These were marked as NA in the HDI variable.
This reduced titles with nation names and HDI scores to
`r nations_no_HDIs` records.
If a place was part of a broader collective (e.g., Scotland or Puerto Rico),
we chose the HDI for the broader political authority
(i.e., the United Kingdom or the United States, respectively).
The HDI includes region level values:
when article titles referenced Africa,
we used the HDI for "Sub-Saharan Africa",
or when article titles referenced Europe,
we used the HDI for "Europe and Central Asia" (*n* = `r regions`).
When multiple place names appeared in the title (*n* = `r multi_nations`),
we averaged the HDI scores.
The UN classifies HDI scores into four categories:
very high (\>= 0.8), high (\>= 0.7), medium (\>= 0.55), and low (\<= 0.549).
The averaged HDI for titles that mention multiple locations
was very high on the HDI scale (*m* = `r multi_HDI_mean`; *mdn* = `r multi_HDI_median`),
indicating that nations with very high HDIs collaborate
with nations that also have very high HDIs.
However, the overall average HDI scores indicate that most nations or places
named in article titles are nations or places with very high HDI scores
(*m* = `r mean_HDI`; *mdn* = `r median_HDI`).

Finally, we added SCImago Journal Rank (SJR) scores
for each of the 61 journal titles in the data set.
We use the SJR to control for citation effects across articles.
SJR scores ranged from
`r min(titles$SJR)` to `r max(titles$SJR)`.
A SJR score below 1.0 indicates below average citations compared to all journals in *Scopus*.
The average SJR in the data was above average
(*m* = `r round(mean(titles$SJR),3)`; *mdn* = `r round(median(titles$SJR),3)`).
However, after deduplicating publication titles,
the average SJR in the data was lower (*m* = `r sjr_unique_mean`; *mdn* = `r srj_unique_median`),
indicating that article titles,
that reference places,
appear more often in journals with higher SJRs.

# Results

This research aimed to investigate the relationship
between the presence of geographical names in journal articles
and their citation counts.
We tested several hypotheses to understand the nuances of this relationship.
Overall, we found that there was a small effect on journal articles
that contained place names in titles.
We were unable to associate this effect with any
of our additional hypotheses.

```{r divide-data, include = FALSE, warning = FALSE}
# data prep
# subset of data that does not include titles with place names
without_nations <- dplyr::filter(titles, is.na(nation))
without_nations <- dplyr::select(without_nations, Citations, Publication, SJR)
# subset of data that does includes titles with place names
with_nations <- dplyr::filter(titles, nation != "NA" & HDI != "NA")
with_nations <- dplyr::select(with_nations, Citations, SJR, Publication, nation, HDI, HDI_Level,
                              CountryRegionMulti, Averaged, places)
```

## Hypotheses

**Main Hypothesis:** Journal articles with geographical names in their titles
are cited less frequently than those without geographical names.

```{r main-hypo, include=FALSE}
citations_wt <- wilcox.test(with_nations$Citations, without_nations$Citations)
ggplot(titles, aes(x = factor(!is.na(nation)), y = log10(Citations + 1))) +
#ggplot(titles, aes(x = factor(!is.na(nation)), y = Citations + 1)) +
  geom_boxplot() +
  xlab("Contains Nation Name in Title") +
  ylab("Citation Count (log_10 + 1)") +
  scale_x_discrete(labels = c("No", "Yes"))
median_with_nations <- median(with_nations$Citations)
median_without_nations <- median(without_nations$Citations)

```

The Mann-Whitney U test revealed a statistically significant difference
in citation distributions between articles with and without
geographical names in their titles.
Further analysis showed that articles with geographical names
had a lower average citaton count
(*mdn*=`r median_with_nations`) than
those without geographical names.
(*mdn*=`r median_without_nations`).
This indicated that article without geographical names
tend to receive slighly more citations.

**Sub-Hypothesis 1:** The closer the geographical name is
to the beginning of the title,
the fewer the citations.

```{r sub-hypo-1, include=FALSE}
cit2places <- cor.test(with_nations$Citations, with_nations$places, method = "spearman")
```

We found a weak but statistically significant `*rho* = r cit2places`
positive correlation between the position
of the geographical name in the title
and the number of citations.

**Sub-Hypothesis 2**: Articles mentioning nations with a higher
Human Development Index in their titles are cited more frequently.

```{r sub-hypo-2, include=FALSE}
cit2HDI <- cor.test(with_nations$Citations, with_nations$HDI, method = "spearman")
```

We found no evidence for a relationship between the number of citations
an article with a geographical name in the title received and the HDI.
The *rho* value near zero (*rho* = `r cit2HDI`)
indicates no meaningful relationship and
the high p-value suggests that this result is likely due to random chance.

**Sub-Hypothesis 3**: The impact of geographical names in titles on citation counts
is moderated by the impact factor of the journal.

```{r sub-hypo-3, include=FALSE}
# rank transform the data
with_nations$Citations_rank <- rank(with_nations$Citations)
with_nations$SJR_rank <- rank(with_nations$SJR)
# create an interaction term
with_nations$interaction_term <- with_nations$places * with_nations$SJR_rank
# kruskal-wallis test
cit2rank <- kruskal.test(Citations_rank ~ interaction_term, data = with_nations)
```

We found that the impact factor of the journal
does not appear to play a moderating role in the relationship
between geographical names in article titles and citation counts,
based on the data.

**Sub-Hypothesis 4:** Journals with a high frequency of articles
with geographical names in titles have lower impact factors.

```{r sub-hypo-4, include=FALSE}
# Group the Data by Journal and Calculate Frequencies
journal_summary <- titles %>%
  group_by(Publication) %>%
  summarise(
    total_articles = n(),
    articles_with_places = sum(!is.na(nation)),
    proportion_with_places = articles_with_places / total_articles
  )
# Merge the Impact Factor Information
journal_summary <- merge(journal_summary, unique(titles[, c("Publication", "SJR")]), by = "Publication")
# Run a Spearman's Rank Correlation Test
cor.test(journal_summary$proportion_with_places, journal_summary$SJR, method = "spearman")
ggplot(journal_summary, aes(x = proportion_with_places, y = SJR)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Proportion of Articles with Place Names") +
  ylab("SJR")

```

We observed a weak, negative correlation between journals
that more frequently publish articles with place names
and their impact factors.
However, given the low p-value based on the data,
this relationship can not be statistically confirmed.

Sub-Hypothesis 5: Are there any trends with where certain nations appear in the titles?

```{r sub-hypo-5, include=FALSE}
# Group Data by Nation and Calculate Average Position
nation_summary <- with_nations %>%
  group_by(nation) %>%
  summarise(
    avg_position = mean(places, na.rm = TRUE),
    total_articles = n()
  )
# correlation test
cor.test(nation_summary$total_articles, nation_summary$avg_position, method = "spearman")
# sort and filter
nation_summary <- nation_summary %>%
  arrange(avg_position) %>%
  dplyr::filter(total_articles >= 10)  # Adjust this threshold as needed
# visualize
ggplot(nation_summary, aes(x = reorder(nation, avg_position), y = avg_position)) +
  geom_bar(stat = "identity") +
  xlab("Nations") +
  ylab("Average Position in Title") +
  coord_flip()
```

While certain nations appear appear more often
in different positions in the article titles,
we found only a weak, negative correlation suggesting
a trend where nations that appear more frequently in titles
may have their names positioned closer to the beginning of the title.

Sub-Hypothesis 6: Journals may be more or less likely to publish articles that mention nations with higher or lower HDIs.

```{r sub-hypo-6, include=FALSE}
# Group Data by Journal and Calculate Average HDI
journal_hdi_summary <- with_nations %>%
  group_by(Publication) %>%
  summarise(
    avg_HDI = mean(HDI, na.rm = TRUE),
    total_articles_with_places = sum(!is.na(places))
  )
# Merge with Total Articles
journal_hdi_summary <- merge(journal_hdi_summary, unique(titles[, c("Publication", "SJR")]), by = "Publication")
# correlation test
cor.test(journal_hdi_summary$avg_HDI, journal_hdi_summary$total_articles_with_places, method = "spearman")
# visualize data
ggplot(journal_hdi_summary, aes(x = avg_HDI, y = total_articles_with_places)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Average HDI of Mentioned Nations") +
  ylab("Total Articles with Place Names")
```

We found a weak, negative correlation that suggested
that journals that publish more articles mentionining place names
might tend to mention nations with slightly lower HDIs,
though this was not statistically confirmed.

# Discussion

Our main hypothesis was that journal articles with
geographical names in their titles were cited less frequently
than those without geographical names,
the difference was statistically significant but minimal.
However, all sub-hypotheses mostly yielded non-significant results,
indicating that the reasons behind this phenomenon are complex
and explained by variables not examined in this study.

Further research might explore other potential factors
that influence this relationship.
These factors may include the specific content of the articles,
the information presented in the abstracts,
the disciplines of the journals, or
broader cultural or academic trends.

Further studies may also examine the citation networks
of articles that include place names.
It could be, that is, that there is a substantial
within-nation network citation advantage to including
place names in articles.
For example, articles that include the geographical name
"Bangladesh" or "United Kingdom" might tend to attract
citations from others in Bangladesh or the United Kingdom, respectively.
Such localized citations could be a reflection of the article's relevance
to that particular context.
Additionally, while including a geographical name might lead
to a perceived loss of generality,
making the article seemingly less applicable to a broader audience,
this disadvantage might be offset by the strength of one's academic, professional, or geographical network.
Researchers from the above listed networks might find the article
more directly relevant to their work and more likely to cite it.

---

This study brought for us a wide set of findings
that addressed the research questions that we set for the study.
Initially, the present study has examined whether
the presence of a country name in the title
reduce the chances of getting more citations.
These findings are discussed.
Papers naming or referring to countries in titles receive fewer citations and
are published in lower ranking journals
than papers that do not name or refer to countries in titles.
In a broad sense,
scientific publications containing
a country name receive lower impact values,
e.g., the impact factor (IF) of the publishing journrals.
However, under the different subject categories
the result can be different [@abramo_effect_2016].
Although Cliff's delta reported negligible effect sizes for citation results,
we think that a cumulative effect might be in play.
That is, if any particular author or journal tends
to publish papers with place names in titles,
then the citation scores for those authors or
journals might receive fewer citations depending
on the frequency of papers with place names in titles.
This result provides general confirmation of our study.
This might be reason that putting the country name
in the title limits the generality of the research findings,
and country-level title is less appealing to the global researchers.
While it is read less,
this leads to get fewer citations.

However, under the countries and HDIs,
this can be different in the context of non-western countries as well.
For example, a researcher from developing countries
has the intuition to read USA-based authors' papers
to know the context and understand the research
in his or her similar research areas.
This helps to increase the chance of developing country
researchers to read good quality papers.
So, if there is a country name in the title and
if it is written by western authors,
it might get the citations by the developing country researchers.
Country names adding at the end of the title,
right part of the title,
do not add much semantic information.
For example, a paper titled
'Information seeking behaviours of the graduate students:
A study on the students from Bangladesh' illustrates a subtitle
does not add much information or value to the article
if researchers prepare the title with the left part of the article.
However, this might be different with some disciplines and articles
where the lengthy titles and occurrences of words in the title matters.
These practices are not fixed and change over times.
While the title is an important factor to the researchers,
there need to be a balance between informativeness and attractiveness.

Although papers that reference country names
in titles received fewer citations and
journals that frequently include articles
that refer to named countries had lower SJR scores,
we could find no relationship between the development level
of a named country with the number of citations.
We could also find no relationship
between the location of the country name in the title
with the article's citations.
Overall, although articles that reference country names
in their titles receive fewer citations and journals have lower SJR scores,
there does not seem to be a geographical bias against specific countries.
We believe that the citation disadvantage associated
with lower citation counts to articles that reference named countries
is primarily a result of perhaps unnecessary specificity
that naming countries adds to the titles and
that makes the articles seem less relevant to others.
While the country name in an article's title refers to less citation and
placed in the lower ranked journals,
we recommend not to use country name in the title.
Even having a country name in the title may lead to bias among the researchers.

## Limitations

It could be that the effects are more pronounced
when sampling from lower impact journals
in some other disciplines.
Here, we only considered the LIS journals.
A future study with journals from other disciplines can add evidence
to support the present result that we got from this study.
Even, the results could be different with a longer citation window.
We also did not control for other factors,
such if a title contains a colon, a question, humor.

# References
